Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-05-02 concurrency/CONCUR 2023/paper.tex
a few more words about deletion

The total lines of code required were about the same as for the original VST verification of the BST. The vast majority of the proof effort was in the templates, in both \lstinline{traverse} and the top-level data structure operations. This is fairly encouraging, since this proof effort should be reusable for other data structures that use the same templates.
%Even though the proof for the \texttt{insert} function within the lock-coupling templates was line-of-code (LoC) intensive, future enhancements could significantly streamline this process. Otherwise, the construction of proofs for the \texttt{traverse} function, in the contexts of both lock-coupling and give-up templates, necessitated a considerable amount of time and effort.
%However, we expect that functions such as \texttt{traverse}, \texttt{insert}, and \texttt{lookup} will be reusable with only minor adjustments needed.  The methods \texttt{findNext} and \texttt{insertOp} may also be adapted to other data structures, such as linked lists, with some additional effort. 
Proving the correctness of e.g. a lock-coupling linked list should be as simple as defining a new \lstinline{node} struct, and then implementing and verifying \lstinline{findNext} and \lstinline{insertOp}.
%Specifically, when attempting to prove concurrent linked lists, we plan to tweak the specifications of \texttt{findNext}, \texttt{inRange}, and \texttt{insertOp} to restructure the \texttt{node} struct. While the size of our template proofs currently approximates that of the no-template versions, we anticipate that strategic application of simple tactics for quantifier expansion could significantly streamline these proofs in the future.
The full development of our mechanization effort is available online at
\href{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}{\color{blue}{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}}.

>>>>

The total lines of code required were about the same as for the original VST verification of the BST. The vast majority of the proof effort was in the templates, in both \lstinline{traverse} and the top-level data structure operations. This is fairly encouraging, since this proof effort should be reusable for other data structures that use the same templates. For lock-coupling deletion, we retained the original (non-templated) proofs; we did not verify give-up deletion, since optimistic deletion is considerably harder to reason about and did not stand to benefit from the template approach.
%Even though the proof for the \texttt{insert} function within the lock-coupling templates was line-of-code (LoC) intensive, future enhancements could significantly streamline this process. Otherwise, the construction of proofs for the \texttt{traverse} function, in the contexts of both lock-coupling and give-up templates, necessitated a considerable amount of time and effort.
%However, we expect that functions such as \texttt{traverse}, \texttt{insert}, and \texttt{lookup} will be reusable with only minor adjustments needed.  The methods \texttt{findNext} and \texttt{insertOp} may also be adapted to other data structures, such as linked lists, with some additional effort. 
Proving the correctness of e.g. a lock-coupling linked list should be as simple as defining a new \lstinline{node} struct, and then implementing and verifying \lstinline{findNext} and \lstinline{insertOp} (plus the appropriate functions for deletion).
%Specifically, when attempting to prove concurrent linked lists, we plan to tweak the specifications of \texttt{findNext}, \texttt{inRange}, and \texttt{insertOp} to restructure the \texttt{node} struct. While the size of our template proofs currently approximates that of the no-template versions, we anticipate that strategic application of simple tactics for quantifier expansion could significantly streamline these proofs in the future.
The full development of our mechanization effort is available online at
\href{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}{\color{blue}{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}}.
